Generate plots to be shown to Coursera users
=====================

This script can be used to generate the pngs shown to users, and also the "data_for_1plots_coursera" file.

Here we're using 9 questions from 8 categories (including the "outlier" category of plots). Each of the 8 cateogries has 10 versions people can see, half are significant half are not. The categories are: small n (n=35), medium n (n=100) (reference), large sample (n=200), best fit lines, axis scale, axis labels, & outlier points.

All p-values are fixed to either be in the range (0.023-0.025) or be in the range (0.33-0.35). We initially neglected the fact that adding an outlier to plots generated in this way (for the "outlier" category) would result in a p-value outside of these ranges. As a result, accuracy rates for "outlier" plots were not directly compared against the reference category in our final analysis (see supplementary materials pdf file).

However, when we planned to use the outlier category, we also decided to remove outliers along the x-asis of all plots, so that only plots in the outlier category would tend to have outlying points. We removed outlying X coordinates by generating 10 extra X coordinates, and then triming off the bottom and top 5.

Saving the plots is currently commented out in this file.


```{r generateBaseData, results='hide', cache=TRUE}
seedForPlots<-234032
set.seed(seedForPlots)

#Intro question:
setwd("/Users/aaronfisher/Documents/JH/EDA Versions/EDA Git Repo/Coursera")

nversions<-5 #how many iterations should we do for each combination of significance and plot category.

#Make vectors to store the p-value bin, category and version number of all the unique questions.
#All p-values will be between .023-.025, or .33-.35 (the outlier points ended up violating this rule).
# We simulate until we get a p-value in the appropriate bin. 
# Note: In the shiny application, which we created afterwards, we used the more elegant solution of binary search to get the appropriate p-value.
pbreaks<-c(.023,.025,.33,.35,1)
pbins.base<-rep(c(2,4),times=8) #used as an index later on for pbreaks variable
nes.base<-c(35,35,100,100,200,200,rep(100,times=10)) #10 points will be added in the generation process and the ten extreme points will be trimmed off
pres.base<-(rep(c('n35','n100ref','n200','bestFit','axesScale','axesLabel','outlier','lowess'),each=2))
pbins<-rep(pbins.base,each=nversions)
nes<-rep(nes.base,each=nversions)
pres<-rep(pres.base,each=nversions)

version.base<-rep(1:nversions) #Note, divide by two because sig and nonsig are two different versions
probnum.base<-c('1-1','1-2','2','3',paste0(rep(4:9,each=2),rep(c('-1','-2'),times=5))) #for labelling plots.
version<-rep(version.base,times=length(pbins.base))  
probnum<-rep(probnum.base,each=nversions) 


cbind(pbreaks[pbins],nes,pres,version)

nreps<-length(pbins)

#Matrixes to hold X coordinates, random errors, and Y coordinates, across all plots.
yes<-matrix(nrow=nreps,ncol=max(nes))
xes<-matrix(nrow=nreps,ncol=max(nes))
ees<-matrix(nrow=nreps,ncol=max(nes))
resids<-matrix(nrow=nreps,ncol=max(nes))

#initialize variables that will store the p-values of each plot
pvals<-rep(0,nreps)
tvals<-rep(0,nreps)
bhat.theory<-rep(0,nreps)
bhat.emp<-rep(0,nreps) #compare the fitted slope coefficient we got with the one we used to generate each plot (bhat.theory)

#First make basic data
#then add outliers

print(paste('nreps = ',nreps))
pb<-txtProgressBar(min = 1, max = nreps,  char = "=", style = 3)  #This progress bar won't show up in the knitr output, as this chunk has it's results hidden.

for(i in 1:nreps){
	tryagain<-TRUE
	while(tryagain){ #repeat until we get a p-value in the desired bin
		n<-nes[i]
		#calculate t-value close to what we want
		t.i<-qt(pbreaks[pbins[i]]/2,df=n-2,lower.tail=F)
		if(abs(t.i)>5) t.i<-0 #guards against when the bin is 1, and quantile has infinite size
		#Add & Trim X coordinates
		x.pre<-rnorm(n +10)
		x<-x.pre[order(x.pre)[6:(length(x.pre)-5)] ] #trim off extra x's
		e<-rnorm(n)
		bhat<-t.i*sd(e)/(sqrt(n)*sd(x)) #sd(e)=true σ, sqrt(n)*sd(x) = Σ[(x-bar(x))^2]
    #In previous iterations of this code, we had it set up to sometimes generate from an actual null (true slope coefficient=0). Below, we disable that option with the added FALSE statement, but we can't get rid of it completely or it will change the seed of the random numbers we generate later on.
		if(FALSE & pbreaks[pbins[i]]>.5 & sample(c(2,2,1),1)==2) {bhat<-0}
		y<-x*bhat*sample(c(-1,1),1)+e
		
		bhat.theory[i]<-bhat
		bhat.emp[i]<-summary(lm(y~x))$coeff[2,1]
		tvals[i]<-summary(lm(y~x))$coeff[2,3]
		pvals[i]<-summary(lm(y~x))$coeff[2,4]
		xes[i,1:n]<-x
		yes[i,1:n]<-y
    ees[i,1:n]<-e
    resids[i,1:n]<-summary(lm(y~x))$resid
    
		pi<-pvals[i]
		bini<-min(which(pi<pbreaks))
		if(bini==pbins[i]) tryagain<-FALSE
		
	}	
	setTxtProgressBar(pb,i) 
}

```



Add an outlier point to each plot in the outlier category

```{r addOutliers, results='hide'}
# For significant plots, outliers are added to the corners.
# For nonsignificant plots, outliers are added above the center of the point cloud.

for(i in which(pres=='outlier')){
  	n<-nes[i]
  	x<-xes[i,1:n]
  	y<-yes[i,1:n]

    sig.i<-pvals[i]<.05
  	#grab the middle point from x and y
  	mx<-rep(mean(x),n)
  	my<-rep(mean(y),n)
  	distvec<-sqrt((mx-x)^2+(my-y)^2)
  	switch<-which(distvec==min(distvec))
  	
    # We put the outlier in one of the upper corners, depending on the slope of the best fit line.
    #case1: #sig slope going down
  	if(tvals[i]<0 & sig.i) { 
  		x[switch]<-min(x)-sd(x)
  		y[switch]<-max(y)+sd(y)
  	}
  	#case2: sig slope going up
  	if(tvals[i]>0 & sig.i) { 
  		x[switch]<-max(x)+sd(x)
  		y[switch]<-max(y)+sd(y)
  	}
  	#case3: flat slope -> move it above the center of the point cloud.
    if(!sig.i){
      y[switch]<-max(y)+sd(y)*sqrt(2) #sqrt 2 makes it the same distance is in cases 1 and 2.
    }
  
  	tvals[i]<-summary(lm(y~x))$coeff[2,3]
  	pvals[i]<-summary(lm(y~x))$coeff[2,4]

	xes[i,1:n]<-x
	yes[i,1:n]<-y

	#NOTE - We do *not* update bhat for the outlier plots
}

#save progress (commented out)
#save(list=c('xes','yes','nreps','pbins','pvals','tvals','nes','pres'),file='data_for_1plots_coursera.RData')
#load('data_for_1plots_coursera.RData')

```

The following code can be used to generate the plot image files shown to users.

```{r createPlots, fig.keep='none'}
#Generate Plots

#To reduce
for(i in 1:nreps){
  n<-nes[i]
  x<-xes[i,1:n]
  y<-yes[i,1:n]
  pval<-pvals[i]
  tval<-tvals[i]
  m<-lm(y~x)
  style<-pres[i]

  t2<-'Data'
  if(style=='lowess') t2<-'with Lowess Line'
  if(style=='bestFit') t2<-'with OLS Best Fit Line'
  title<-paste('Sample ',t2,sep='')
  xl<-paste("Cranial Electrode",floor(runif(1,11,44)),"(Standardized)")
  yl<-paste("Cranial Electrode",floor(runif(1,53,97)),"(Standardized)")
  
  #Saving plots is commented out
  #png(paste0("pvalue plot images/coursera2_#",probnum[i],'_datVer-',version[i],'_',pres[i],'_pval-',round(pvals[i],digits=3),".png"), width = 400, height = 400)
    par(mfrow=c(1,1))
  	plot(x,y,xlab='X',ylab='Y',main=title)
  	if(style=='lowess') lines(lowess(x,y))
  	if(style=='bestFit') abline(m$coef)
  	if(style=='axesScale')plot(x,y,xlab='X',ylab='Y',main=title,xlim=c(min(x)-1.5*sd(x),max(x)+1.5*sd(x)),ylim=c(min(y)-1.5*sd(y),max(y)+1.5*sd(y)))
 
  	if(style=='axesLabel') plot(x,y,xlab=xl,ylab=yl,main=title)
  #dev.off()
}

```


